---
title: "Airbnb in New York City"
author: "Carole Mattmann und Jonas Zuercher"
date: "27 February 2020"
output:
  html_document: default
  pdf_document: default
  word_document: default
always_allow_html: true
---
# Summary
We are exploring a dataset of airbnb listings in New York City in 2019. 

Analyses on the prices of the listings were run and models were created to predict prices of the listings. The best model to calculate the price that was found includes 5 variables:

* room type (factor with 3 levels, entire appartment being the highest and shared room the lowest)
* distance to timessquare (negative effect)
* availability (positive effect)
* neighbourhood group (factor with 5 levels, Manhattan being the highest and Bronx the lowest)
* minimum nights (negative effect)

The airbnb dataset was merged with a dataset concerning incidents (e.g. crimes) in the concerning neigbourhoods. 

The airbnb dataset is vizualized in two maps in the last chapter in two ways:

* All  datapoints are visualized on the map and coloured depending on their price range.
* The most expensive and cheapest housings are visualized with personalized markers. Also an distance calculator is included in the map. 

# Packages 

We used the following packages to work with the data. 

```{r packages, include=TRUE, echo=TRUE, eval = TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyverse)
library(geosphere)
library(ggplot2)
library(tinytex)
library(leaflet)
```
# Data import and cleaning

## Airbnb dataset

The dataset was downloaded from: https://www.kaggle.com/dgomonov/new-york-city-airbnb-open-data

### Import
```{r import, eval=TRUE, echo=TRUE, include=TRUE}
AB_NYC <- read.csv("../01_data/AB_NYC_2019.csv", header=TRUE)

```

### Overview of dataset
```{r, eval=TRUE, echo=TRUE, include=TRUE}
str(AB_NYC,width=80,strict.width="cut")

```

Following changes are made to the dataset:

### Remove price 0

Remove all listings with price 0

```{r price0, include=TRUE, echo=TRUE, eval = TRUE}
AB_NYC <-AB_NYC[AB_NYC$price > 0,]
```

### Add log price

Add logarithmic price for analysis purposes 
```{r pricelog, include=TRUE, echo=TRUE, eval = TRUE}
AB_NYC <- cbind(AB_NYC,price_log = log(AB_NYC$price))
```

### Remove inactive listings

Remove inactive listings and make new dataset 

```{r removeinactive, include=TRUE, echo=TRUE, eval = TRUE}
AB_NYC_available <- AB_NYC %>% 
  filter(availability_365 > 0) 
```

### Add distance to Times Square to model

We want to make a statement about how central the place is. Therefore the distance to Times Square is caculated using the latitude and longitude of the listings. The package "geosphere" is used.

Times Square, Manhattan, NY, USA, Latitude and longitude coordinates are: 40.758896, -73.98513

```{r distance_timessquare, include=TRUE, echo=TRUE, eval = TRUE}

coord <- cbind(AB_NYC_available$longitude,AB_NYC_available$latitude)
dist.timessquare <- distGeo(p1=coord, p2=c(-73.985130, 40.758896))
AB_NYC_available <- cbind(AB_NYC_available,dist.timessquare)
```

### Prepare dataset for merging with the second dataset

We want to merge this dataset with the Incident dataset. Therefore we need to transform the neighbour_group column by changing the letters to lowercase, delete empty spaces and defining the entries as factors. 

```{r echo=TRUE}
# write neighbourhood group entries in lower case
AB_NYC_available$neighbourhood_group<-tolower(AB_NYC_available$neighbourhood_group)

# remove spaces from neighbourhood groups
AB_NYC_available$neighbourhood_group <-gsub(" ","", AB_NYC_available$neighbourhood_group)

# neighbourhood group as factor
AB_NYC_available$neighbourhood_group<-factor(AB_NYC_available$neighbourhood_group)
```

## Incidents dataset
The dataset was downloaded from: https://data.cityofnewyork.us/City-Government/Agency-Performance-Mapping-Indicators-Annual/gsj6-6rwm 

```{r}
Ind_NYC<- read.csv("../01_data/Indicators_NYC.csv")

```
### Overview of dataset
```{r}
str(Ind_NYC,width=80,strict.width="cut")
```

### Produce a dataframe for 2019 including Neighbourhoods, indicators and the incident numbers

First a dataframe which only contains the neighbour_group, the indicator and the incidents of 2019 has been produced. 

```{r}
Ind_NYC_2019_cleaned<-data.frame("neighbourhood_group"= Ind_NYC$Geographic.Identifier,
                                 "Indicator"=Ind_NYC$Indicator,
                                 "Incidents"=Ind_NYC$FY2019)
head(Ind_NYC_2019_cleaned)
```
### Transform dataframe for further processing

We want to merge this dataframe with the airbnb dataset. Therefore we need to transform the neighbour_group column by removing the numbers of the neighbourhood_group entries, changing the letters to lowercase, delete empty spaces and defining the entries as factors. 

```{r}
#Remove numbers from neighbourhood_group column
Ind_NYC_2019_cleaned$neighbourhood_group <-gsub("[0-9]","", Ind_NYC_2019_cleaned$neighbourhood_group)
#Remove empty spaces from neighbourhood_group column
Ind_NYC_2019_cleaned$neighbourhood_group <-gsub(" ","", Ind_NYC_2019_cleaned$neighbourhood_group)
#Change all lettes from neighbourhood_group column to lower case
Ind_NYC_2019_cleaned$neighbourhood_group<-tolower(Ind_NYC_2019_cleaned$neighbourhood_group)
#Change type of the neighbourhood_group column to factor
Ind_NYC_2019_cleaned$neighbourhood_group<-factor(Ind_NYC_2019_cleaned$neighbourhood_group)

#overview
head(Ind_NYC_2019_cleaned)

```
### Summarise Observation per neighbourhood_group

We want only one row per neighbourhood_group entry and Indicator. Therefore we first group the entries by neighbourhood_groups and then summarize the incidents numbers. Because some entries are given in percentage those entries result in NA entries after this process and need to be removed with the filter feature. 
```{r}
# sum of incidents per neighbourhood group and indicator 
Summary_Ind_NYC_2019<-Ind_NYC_2019_cleaned %>%
  group_by(neighbourhood_group=Ind_NYC_2019_cleaned$neighbourhood_group,Indicator) %>%
  summarise(Observations=sum(Incidents,na.rm = TRUE))

# Remove entries without a neighbourhood group from the dataframe
Summary_Ind_NYC_2019<-filter(Summary_Ind_NYC_2019,neighbourhood_group != "")
head(Summary_Ind_NYC_2019)

```
### Transform dataframe into a nested dataframe

The dataframe is transformed into a nested dataframe to improve readiablity after merging it with the airbnb dataset. 

```{r}
# nested indicators
NYC_nest<-Summary_Ind_NYC_2019 %>%
  nest(Indicator=c(Indicator, Observations))
head(NYC_nest)
```

## Merge datasets

The processed dataset are merged together by the neighbour_group column. 

```{r message=FALSE, warning=FALSE}
#Join both datasets
NYC<-left_join(AB_NYC_available,NYC_nest, by="neighbourhood_group")

# neighbourhood group as factor
NYC$neighbourhood_group<-factor(NYC$neighbourhood_group)
```


# Data visualisation

## Distribution of prices by room types and neighbourhood

Prices of airbnb listings are shown in boxplots, grouped by room type and location (neighbourhood group) are shown below.

```{r plot1, include=TRUE, echo=TRUE, eval = TRUE}

ggplot(data = AB_NYC_available,
       mapping = aes(y = price_log,
                     x = "",
                     group = neighbourhood_group,
                     colour = neighbourhood_group)) +
  geom_boxplot() +
  facet_wrap(. ~ room_type)+
  xlab("")+
  ylab("log (price in $ per night)")

```

Prices of the room type "entire home/apt" have the highest median, followed by "private room"" and lastly "shared room"", which is not surprising. 25. and 75. quantile for "entire home/apt" and "private room" are similarly distributed, for shared room there is no clear pattern. 

For all room types, median prices in neighbourhood "Manhattan" are the highest. For for "entire home/apt" and "private room" the second highest mediam prices are in Manhattan. 

## Distribution of prices 

The distribution of prices, grouped by room type is shown below.

```{r, include=TRUE, echo=TRUE, eval = TRUE}

ggplot(data = AB_NYC,
       mapping = aes(x = price_log,
                     group = room_type,
                     colour = room_type,
                     fill = room_type,
                     alpha = 0.5)) +
  geom_density() +
  xlab("log (price in $ per night)")+
  ylab("density ")
```

Prices for all room types are skewed to the right. Even with logarithmic display of prices, this is still clearly the case. 

## Availability of appartments 

The availability of the listings is shown below in days per year.

```{r, include=TRUE, echo=TRUE, eval = TRUE, warning=FALSE, message=FALSE}

ggplot(data = AB_NYC_available,
       mapping = aes(x = availability_365)) +
  geom_histogram() +
  xlab("availability in days per year")+
  ylab("number of listings")
```

There are a lot of listings with very low or very high (almost year round) availablity. Listings with no available days in 2019 were removed from the dataset. This distribution was not taken into account when looking at the prices.  


# Possible models to calculate the price of an airbnb

## Simple linear models

Impact of several variables on the price are analysed. The highest R2 is reached with "room type".

```{r, include=TRUE, echo=TRUE, eval = TRUE}
# price ~ neighbourhood group

lm.hood <- lm (data=AB_NYC_available, price_log~neighbourhood_group)
summary(lm.hood)

# price ~ room type

lm.type <- lm (data=AB_NYC_available, price_log~room_type)
summary(lm.type)

# price ~ dist.timessquare

lm.dist <- lm (data=AB_NYC_available, price_log~dist.timessquare)
summary(lm.dist)
```

## Multiple linear model

A linear model using "distance to timessquare" and "room type" (as factor) with an assumed interaction is applied to calculate the price.

```{r, include=TRUE, echo=TRUE, eval = TRUE, warning=FALSE}
#distance and room type on price (with interaction)
lm.dist.type.interact <- lm (data=AB_NYC_available, price_log~dist.timessquare*room_type)
summary(lm.dist.type.interact)

```

The model is plotted using three different regression lines per room type. There is a tendency for all room types that the price is lower if the place is further from Times Square. 

Please note: The interactions are not significant considering the p-values.  

```{r, include=TRUE, echo=TRUE, eval = TRUE, warning=FALSE}
ggplot(data = AB_NYC_available,
       mapping = aes(y = log(price),
                     x = dist.timessquare,
                     colour = room_type,
                     group = room_type)) +
  geom_point(alpha = 0.03) +
  xlab("distance to timessquare in m")+
  ylab("log (price in $ per night)")+
  geom_smooth(method="lm")
```


## Multiple linear model by choosing smallest RSS

A multiple linear model is created. The best model to calculate the price we can find includes 5 variables:

* room type (factor with 3 levels, entire appartment being the highest and shared room the lowest)
* distance to timessquare (negative effect)
* availability (positive effect)
* neighbourhood group (factor with 5 levels, Manhattan being the highest and Bronx the lowest)
* minimum nights (negative effect)


```{r lm2, eval=TRUE, include=TRUE, echo=TRUE, results=FALSE}

#full model
lm.full <- lm (data=AB_NYC_available, price_log~room_type
               +neighbourhood_group
               +minimum_nights
               +number_of_reviews
               +calculated_host_listings_count
               +availability_365
               +dist.timessquare)

#empty model
lm.empty <- lm (data=AB_NYC_available, price_log~NULL)
add1(lm.empty,scope=lm.full)

#choose value with smallest RSS
lm.1 <- update(lm.empty,.~.+room_type) 
add1(lm.1,scope=lm.full)
lm.2 <- update(lm.1,.~.+dist.timessquare)
add1(lm.2,scope=lm.full)
lm.3 <- update(lm.2,.~.+availability_365)
add1(lm.3,scope=lm.full)
lm.4 <- update(lm.3,.~.+neighbourhood_group)
add1(lm.4,scope=lm.full)
lm.5 <- update(lm.4,.~.+minimum_nights)
add1(lm.5,scope=lm.full)
```

```{r, include=TRUE, echo=TRUE, eval = TRUE}
summary(lm.5)
```



# Interactive maps with the leaflet package

In this part the leaflet package is used to produced interactive maps of the merged dataset. 

## Interactive map with three coloured price groups

The first interactive map shows all housings with a colored marker. The color varies with the price per night for the housings. 

### Prepare groups for visualization in the map

To adjust the color depending on the price of the datapoints a if-else loop is used. The same approach is used to classify the datapoint into the three groups Cheap, Medium and Expensive. 

```{r}
# Function that chose the color depending on the price of the housing
getColor <- function(NYC) {
  sapply(NYC$price, function(price) {
    if(price <= 50) {
      "green"
    } else if(price <= 200) {
      "orange"
    } else {
      "red"
    } })
}

# Function that chose a group name depending on the price of the housing
getGroup <- function(NYC) {
  sapply(NYC$price, function(price) {
    if(price <= 50) {
      "Cheap"
    } else if(price <= 200) {
      "Medium"
    } else {
      "Expensive"
    } })
}
```

### Visualization of a color coded interactive map

Its possible to zoom and change settings in this interactive map. Two different base map layers can be chosen. Furthermore only certain price groups like Cheap can be shown in the map. The green points are housings up to 50 dollar per night. The yellow points are housings up to 200 $ per night and the red ones are the housing that are even more expensive. Each marker shows detailed information with a popup if its clicked.  

```{r, message=FALSE, warning=FALSE}
map<-leaflet(NYC) %>%
  #Base layer
  addTiles(group = "OSM") %>%
  addProviderTiles(providers$CartoDB.Positron,group="GREY")%>%
  
  # Overlay groups - Adding price dependant colored housing markers
  addCircleMarkers(lng = NYC$longitude, 
                   lat = NYC$latitude, 
                   color=getColor(NYC),
                   popup = paste0("Price per night: ", as.character(NYC$price), " $",
                                         "<br>","Room type: ",NYC$room_type, 
                                         "<br>","Distance to Times Square: ",
                                  as.character(round(NYC$dist.timessquare/1000, digits=1)), " km"),
                   radius=0.1,group = getGroup(NYC)) %>%
  
  #LayercOntrol
  addLayersControl(baseGroups = c("OSM","GREY"), 
                   overlayGroups = c("Cheap","Medium","Expensive"))


map
```

## Interactive map with the most expensive and cheapest housings 

The second interactive map shows the most expensive and the cheapest housings of the dataset with personalized markers. 

### Prepare dataframe for cheapest and most expensive spots in New York

First the most expensive and the cheapest housings have to be found in the dataset and saved in a dataframe. 

```{r echo=FALSE}
# Most expensive place in the NYC dataframe
df_exp<-filter(NYC,price == max(price))

# Cheapest places in the NYC dataframe
df_cheap<-filter(NYC,price == min(price))
```
### Prepare Icons
The icons have to be prepared for the visualization. Therefore the pictures are imported. 
```{r}
castle<-makeIcon(
  iconUrl = "../01_data/castle.png",
  iconWidth = 20, iconHeight = 20)

hovel<-makeIcon(
  iconUrl = "../01_data/hovel.png",
  iconWidth = 20, iconHeight = 20)

```

### Visualization of the most expensive and cheapest airbnb housings in New York

Its possible to zoom and change settings in this interactive map. Two different base map layers can be chosen. Furthermore only the cheapest or the most expensive datapoints can be shown in the map. The most expensive housing is visualized with a custom castle icon. The cheapest housings are visualized with a custom hovel icon. Each marker shows detailed information with a popup if its clicked. Furthermore distances between two points can be measured with the measurement tool implemented in the map.  

```{r}
map2<-leaflet(NYC) %>%
  #Base
  addTiles(group = "OSM") %>%
  addProviderTiles(providers$CartoDB.Positron,group="GREY")%>%
 
  # Measure
 
  addMeasure(primaryLengthUnit = "kilometers",activeColor="red")%>%
  
  # Overlay groups
  addMarkers(lng=df_exp$longitude, 
             lat=df_exp$latitude, 
              icon= castle, popup = paste0("Price per night: ", 
                                           as.character(df_cheap$exp), " $","<br>","Room type: ",
                                         df_exp$room_type, "<br>","Distance to Times Square: ",
                                         as.character(round(df_exp$dist.timessquare/1000, digits=1))," km"),
             group=("Most expensive Airbnb Housing")) %>%
  addMarkers(lng=df_cheap$longitude,
             lat=df_cheap$latitude, 
             icon= hovel, popup = paste0("Price per night: ", 
                                         as.character(df_cheap$price)," $", "<br>","Room type: ",
                                         df_cheap$room_type, "<br>","Distance to Times Square: ",
                                         as.character(round(df_cheap$dist.timessquare/1000,
                                                            digits=1))," km"),
             group=("Cheapest Airbnb Housing")) %>%
  #LayerCOntrol
  addLayersControl(baseGroups = c("OSM","GREY"), 
                   overlayGroups = c("Most expensive Airbnb Housing","Cheapest Airbnb Housing"))

map2
```



